---
title: "answer_key"
output: html_document
---

## Try with some real data
Your turn. See if you can:
1. Read in the file called `trapping_data.csv` 

```{r}
trap_data <- read_csv(here("Data/trapping_data.csv"))
```

2. Inspect the data. Is it tidy? Write yourself some notes about your answer.

The data are tidy. There is one row per observation and one observation per row. Hints that it is a tidy format include the fact that in the Habitat and "Ground_or_Tree" columns, the same information is repeated multiple times, for each observation.


3. Cast the data so that instead of a column called "Habitat" there are three columns called "Pine", "Deciduous" and "Hemlock". Put the values for "Animal_wt_g" in those new columns.

Here's how we do that without saving a new object, using something called "piping":

```{r}
trap_data |> pivot_wider(
  names_from = Habitat, #make 3 new columns using names from Habitat column
  values_from = Animal_wt_g) |> #spread values from Animal_wt_g into the 3 new columns
  View() #open a view of the pivoted data
```

4. Comment your code appropriately as you work.
5. Make a few commits as you work.

You will use:
- `git add practice-tidy-git.Rmd`
- `git commit -m "my commit message" `
in a desktop client or in the RStudio terminal window

6. Push your changes to your github repo.

You will use:
- `git push` OR
- `git push origin main`
 in a desktop client or in the RStudio terminal window
