---
title: "answer_key"
output: html_document
---

## Try with some real data
Your turn. See if you can:

1. Read in the file called `trapping_data.csv` 
Need to make sure you have the proper packages turned on.

```{r}
library(tidyverse) #for tidying data
library(here) #for helping with file paths
```


```{r}
trap_data <- read_csv(here("Data/trapping_data.csv"))
```

2. Inspect the data. Is it tidy? Write yourself some notes about your answer.

The data are tidy. There is one row per observation and one observation per row. Hints that it is a tidy format include the fact that in the Habitat and "Ground_or_Tree" columns, the same information is repeated multiple times, for each observation.


3. Cast the data so that instead of a column called "Habitat" there are three columns called "Pine", "Deciduous" and "Hemlock". Put the values for "Animal_wt_g" in those new columns.

Here's how we do that without saving a new object, using something called "piping":

```{r}
trap_data |> pivot_wider(
  names_from = Habitat, #make 3 new columns using names from Habitat column
  values_from = Animal_wt_g) |> #spread values from Animal_wt_g into the 3 new columns
  View() #open a view of the pivoted data
```

4. Comment your code appropriately as you work.
5. Make a few commits as you work.

You will use:
- `git add practice-tidy-git.Rmd`
- `git commit -m "my commit message" `
in a desktop client or in the RStudio terminal window

6. Push your changes to your github repo.

You will use:
- `git push` OR
- `git push origin main`
 in a desktop client or in the RStudio terminal window
 
7. Read in the file called `trapping_data1.csv` 

```{r}
trap_data2 <- read_csv(here("Data/trapping_data1.csv"))
```


8. Inspect the data. Is it tidy? Write yourself some notes about your answer.

These data are less tidy than they could be. While Groud_or_Tree seems to be one observation per row, the 3 columnd Pine, Hemlock, and Deciduous are likely intended to describe the habitat, and could be more tidy if they were recast into a longer format.

9. Cast the data so that instead of columns called "Pine", "Deciduous" and "Hemlock" there is one column called "Habitat" Put the values from those 3 columns into a new column called "Animal_wt_g".

This time, we'll save the recast data as a new object called `new_data`
```{r}
new_data <- pivot_longer(
  trap_data2, #dataset to be pivoted
  cols = Pine:Deciduous, #identify the columns that need to be cast
  names_to = "Habitat", #identify the name for the new column that will take all the course names
  values_to = "Animal_wt_g") #identify the name for the column in which the values will be stored)
```

Notice that it works, but not as nicely as if you had set the data up in a tidy format to begin with in your spreadsheet software.

10. Comment your code appropriately as you work.

11. Make a few commits as you work.
You will use:
- `git add practice-tidy-git.Rmd`
- `git commit -m "my commit message" `
in a desktop client or in the RStudio terminal window

12. Push your changes to your github repo.
You will use:
- `git push` OR
- `git push origin main`
 in a desktop client or in the RStudio terminal window



