---
title: "practice-tidy-git"
output: html_document
---

## Let's apply what we've learned!

Let's set up the workspace

```{r, include=FALSE}
rm(list = ls()) #clears RStudio environment for a fresh start
library(tidyverse) #turns on all tidyverse packages; must be installed with install.packages("tidyverse) first time
library(here) #turns on here "pointer" package; must be installed with install.packages("here") first time
```

Let's read in some data sets

```{r}
not_tidy1 <- read_csv(here("Data/not_tidy1.csv"))
not_tidy2 <- read_csv(here("Data/not_tidy2.csv"))
not_tidy3 <- read_csv(here("Data/not_tidy3.csv"))
our_tidy_data <- read_csv(here("Data/our_tidy_data.csv"))
```

## Interpret
What problems do you see in `not_tidy1`? Write them here:





\
\
After you write the problems you identify in lines 27 - 31, use git to **add** and **commit** your changes!
Then try to **push** your changes, and check to see that they are transferred to your github repo!

\

What problems do you see in `not_tidy2`? Write them here:





\
\
After you write the problems you identify in lines 40-44, use git to **add** and **commit** your changes!
Do you notice any changes in your github repo?

\
\
What problems do you see in `not_tidy3`? Write them here:





\
\

After you write the problems you identify in lines 53-57, use git to **add** and **commit** your changes!
Then try to **push** your changes, and check to see that they are transferred to your github repo!
\

## How do I fix it?
We see that both `not_tidy2` and `not_tidy3` have all of the same information as
`our_tidy_data` but not the same structure.

We call `our_tidy_data` a "long" format, vs. the "wide" format of `not_tidy2` and `not_tidy3`.

Fortunately, it is easy to move back and forth between long and wide format.
So, if someone gave you data formatted as `not_tidy_2` or `not_tidy3` you could rearrange it from wide to long format.

Sometimes, a particular statistical test also requires your data to be in a wide format. 

Let's look at how to do it.

### Change `not_tidy2` from wide to long format

We use a function called `pivot_longer` from the `tidyr` package (`tidyr` is installed automatically with `library(tidyverse)`).

Here's how it works:

```{r}
not_tidy2_long <- pivot_longer(
  not_tidy2, #dataset to be pivoted
  cols = Mammalogy:`English literature`, #identify the columns that need to be cast
  names_to = "Course", #identify the name for the new column that will take all the course names
  values_to = "Grade") #identify the name for the column in which the values will be stored)
```

Now see if you can cast `not_tidy_3` from wide to long into an object called `not_tidy_3_long`:



What if we wanted to cast from long to wide? The process is similar.

Let's try it!

```{r}
our_tidy_wide <- pivot_wider(
  our_tidy_data, #data set name
  names_from = Course, #variable that will be cast into more columns
  values_from = Grade) #variable from which to take the values to put in the new columns
```

Starting to make sense?

## Try with some real data
Your turn. See if you can:
1. Read in the file called `trapping_data.csv`
2. Inspect the data. Is it tidy? Write yourself some notes about your answer.
3.Cast the data so that instead of a columns called "Habitat", you have 3 new columns called "Pine", "Deciduous" and "Hemlock". Put the values from the column called "Animal_wt_g" into those 3 new columns.
4. Comment your code appropriately as you work.
5. Make a few commits as you work.
6. Push your changes to your github repo.
7. Read in the file called `trapping_data1.csv` 
8. Inspect the data. Is it tidy? Write yourself some notes about your answer.
9. Cast the data so that instead of columns called "Pine", "Deciduous" and "Hemlock" there is one column called "Habitat" Put the values from those 3 columns into a new column called "Animal_wt_g".
10. Comment your code appropriately as you work.
11. Make a few commits as you work.
12. Push your changes to your github repo.
13. Solutions are in the .Rmd file called `answer_key.Rmd` in the Code folder.
